---
title: "Specific Stage Analysis"
author: "Matthew J Cashman"
date: "September 30, 2018"
output:
  html_document: null
  pdf_document: default
  highlight: tango
  theme: united
  toc: yes
  toc_float: yes
---

# Pre-analysis
## Parameter selection
Before we begin with our specific stage analysis, there are several settings that we will have to define first. 1. Site  - USGS site number to do the analysis on. 2. startDate - Beginning of the period of interest, if available. Blank will pull from the earliest record. 3. endDate - End of the period of interest. Blank will pull from the latest record. 4. Quantiles - Mean daily flow quantiles, in decimal form, for the analysis. e.g. 0.8 is the 80% of mean daily flow, or 20% exceedance.5. Sensitivity - Used in the clean up procedure. Will reject any values outside of discharge quantile equation: +/- sensitivity, in percent 6. target - Target quantile of interest, used to create single, focused output graphs 

```{r Setting analysis parameters}
Site <- "01589035"
startDate <- "2010-10-29"
endDate <- "2017-09-31"
Quantiles <- c(0.3,0.5,0.7,0.8,0.85,0.9,0.95,0.98,0.99,0.9975,0.999,0.9999,0.9999999)
Sensitivity <- 0.01
```



```{r Loading Required Packages, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(tidyverse, ggthemes, fs, stringr, dataRetrieval, data.table,tictoc, microbenchmark,fasttime)
```

```{r Define Functions, echo = FALSE}
import_sites <- function(Site, startDate, endDate) {
  message("Pulling unit value data from NWIS might take some time, please be patient")
  message("Reading Site Info")
    sitedata <<- readNWISsite(siteNumbers = Site)
  message("Reading Daily Data")
    daily <<- readNWISdv(siteNumbers = Site,
                      parameterCd = c("00060"),
                      startDate = startDate,
                      endDate = endDate) %>%
                      renameNWISColumns() %>%
                      filter(Flow!=-999999)
  message("Reading Continuous Data")
    uvdata <<- readNWISuv(siteNumbers = Site,
                       parameterCd = c("00060","00065"),
                       startDate = startDate,
                       endDate = endDate) %>% 
                       renameNWISColumns() %>%
                       filter(Flow_Inst!=-999999) %>%
                       filter(GH_Inst!=-999999)
}
write_sites <- function(Site) {
  writeDir <- paste0("./NWIS_pulls/",Site)
  dir_create(writeDir) #Create Export directory
  message("Saving Daily Data locally")
  fwrite(daily,paste0(writeDir,"/daily.csv"))
  message("Saving Continuous Data locally")
  fwrite(uvdata,paste0(writeDir,"/uvdata.csv"))
  message("Saving Site Info locally")
  fwrite(sitedata,paste0(writeDir,"/sitedata.csv"))
}
read_local <- function() {
  readDir <- paste0("./NWIS_pulls/",Site)
  message("Reading Daily Data locally")
    daily <<- fread(paste0(readDir,"/daily.csv"), colClasses=c(site_no="character"))
    daily$Date <<- as.Date(daily$Date)
  message("Reading Continuous Data locally")
    uvdata <<- fread(paste0(readDir,"/uvdata.csv"), colClasses=c(site_no="character"))
    uvdata$dateTime <<- fastPOSIXct(uvdata$dateTime)
  message("Reading Site Info locally")
    sitedata <<- fread(paste0(readDir,"/sitedata.csv"), colClasses=c(site_no="character"))
}
theme_ss <- function(base_size, base_family) {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family=base_family)
    + theme(plot.title = element_text(size = rel(1), hjust = 0.5),
            plot.subtitle = element_text(size = rel(01), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2, size = rel(0.9)),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size=rel(0.8)), 
            axis.text.x = element_text(size=rel(0.8)), 
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "right",
            # legend.direction = "verticale",
            # legend.key.size= unit(0.5, "cm"),
           # legend.margin = unit(0, "cm"),
            legend.title = element_text(size = rel(0.8)),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text()
    ))
}
SS <- function(Site, Quantiles) {
  quantiles <<- quantile(daily$Flow, probs=Quantiles, na.rm=TRUE)
  Runlist = list()
  for(i in 1:(length(quantiles))) {
    Discharge <- quantiles[[i]]
    # Find points where Flow_Inst is above target discharge
    # Points always intersect when above=TRUE, then FALSE or reverse 
    above<-uvdata$Flow_Inst>Discharge
    intersect.index<-which(diff(above)!=0) #Vector of index values immediately before flow passes quantile
    
    # Find the slopes for each line segment.
    Intersect.slopes <- uvdata$Flow_Inst[intersect.index+1]-uvdata$Flow_Inst[intersect.index]
    # Estimate time where target quantile crosses interpolated discharge
    x.points <- intersect.index + ((Discharge - uvdata$Flow_Inst[intersect.index]) / (Intersect.slopes))
    intersect.points <- as.data.frame(x.points)
    
    suppressWarnings(index.fraction <- intersect.points %>% #Split decimals
      separate(x.points, into = c("index","fraction")))
    
    index.fraction$fraction <- index.fraction$fraction %>% #Replace NA with 0's
      replace_na(0)
    
    index.fraction$fraction <- paste("0.", index.fraction$fraction, sep="") #Add 0. to front of decimal values
    index.fraction$secondspast <- round(as.numeric(index.fraction$fraction)*15*60)
    
    QestTime<-uvdata$dateTime[as.numeric(index.fraction$index)]+index.fraction$secondspast #Estimated time of target Q
    
    Approx_Flow <- as.tibble(approx(uvdata$dateTime, y = uvdata$Flow_Inst, xout = QestTime))  %>%
      left_join(as.tibble(approx(uvdata$dateTime, y = uvdata$GH_Inst, xout = QestTime)), by = "x") %>%
      rename(Flow_Inst = y.x) %>%
      rename(GH_Inst = y.y) %>%
      rename(dateTime = x)
    Approx_Flow$i <- paste(i)
    Approx_Flow$Quantflow <- quantiles[[i]]
    Runlist[[i]] <- Approx_Flow
  }
  
  #Bind quantile lists into one dataframe
  SS_results <<- bind_rows(Runlist)
  
  #Cleanup <- not needed anymore due to global binding in function?
  #rm(Approx_Flow,index.fraction,intersect.points,Runlist)
}

```

##Load NWIS data
Here we are loading data that has been loaded locally. Options exist to pull directly from NWIS using the dataRetrieval package. Error values or missing data (e.g. values of -999999) are excluded at this point.
```{r Import/Read Data, echo = FALSE}
 #import_sites(Site, startDate, endDate)
 #write_sites(Site)
  #OR#
  read_local()
```

##Plot NWIS data
After our data has been downloaded, we will plot mean daily discharge, instantaneous unit discharge, and instantaneous unit stage to check for any gaps or weird values. Directly measured values are indicated by red points.

```{r Plot Loaded DV/UV Data, echo = FALSE}
meas<-readNWISmeas(siteNumber = Site) %>%
  select(party_nm,gage_height_va,discharge_va,gage_va_change,gage_va_time,discharge_cd,measurement_dateTime,tz_cd)

#Plot daily data
ggplot(data=subset(daily), aes(x=Date, y=Flow))+
    ggtitle(paste("USGS ", Site, sitedata$station_nm), subtitle = "Mean Daily Hydrograph")+
    theme_ss(14,"sans")+
    geom_line(alpha=0.8)+
    ylab("Mean Daily Discharge (cfs)")

#Plot unit value discharge & stage
ggplot(data=subset(uvdata), aes(x=dateTime, y=Flow_Inst))+
    ggtitle(paste("USGS ", Site, sitedata$station_nm), subtitle = "Instantaneous Unit Value Hydrograph")+
    theme_ss(14,"sans")+
    geom_line(alpha=0.8)+
    ylab("Instantaneous Discharge (cfs)")+
     geom_point(data=meas, aes(x=measurement_dateTime, y=discharge_va), color="red")


ggplot(data=subset(uvdata), aes(x=dateTime, y=GH_Inst))+
  ggtitle(paste("USGS ", Site, sitedata$station_nm), subtitle = "Instantaneous Unit Value Stage")+
  theme_ss(14,"sans")+
    geom_line(alpha=0.8)+
    ylab("Gage Height (ft)")+
     geom_point(data=meas, aes(x=measurement_dateTime, y=gage_height_va), color="red")

```

#Specific Stage Analysis
##Run the Analysis
```{r Run Specific Stage Analysis}
#Run Specific Stage Analysis ----
SS_results <- SS(Site, Quantiles)
```

##Cleanup
Now we check results for stage and discharge to make sure there are no outliers. In the figure below, we can see some scatter around our target discharge quantiles.

```{r Plot Check for Outliers, echo=FALSE}
ggplot(data = SS_results, aes(x=dateTime, y=Flow_Inst, color=as.factor(i)))+
  geom_point(alpha=0.5)+
  theme_ss(14, "sans")+
  ggtitle(paste("USGS ", Site, sitedata$station_nm), subtitle = "Discharge Quantiles Check")+
  guides(colour = guide_legend(override.aes = list(alpha = 1),reverse=TRUE))+
  scale_color_discrete(name="Discharge\nQuantile", breaks = unique(SS_results$i), labels=(paste(colnames(quantiles),":",round(quantiles,0)," cfs")))+
  scale_x_datetime(date_labels = "%b\n%Y")+
  ylab("UV Discharge (cfs)")
```

A simple dplyr pipe uses mutate to calculate the percent difference between the calculated discharge and the discharge of interest. Results are then filtered to select only those values less than the percent difference specified by Sensitivity.
```{r Clean up Outliers, warning=FALSE}
SS_results_clean <- SS_results %>%
                    mutate(pc_diff = (Flow_Inst-Quantflow)/Quantflow) %>%
                    filter(pc_diff > -Sensitivity & pc_diff < Sensitivity)

```

## Post-outlier Check
Now we check results again to make sure all outliers have been removed.
```{r Plot Cleaned results, echo=FALSE, message=FALSE,warning=FALSE}
ggplot(data = SS_results_clean, aes(x=dateTime, y=Flow_Inst,color=as.factor(i)))+
  geom_point(alpha=0.5)+
  theme_ss(14, "sans")+
  ggtitle(paste("USGS ", Site, sitedata$station_nm), subtitle = "Discharge Quantiles Check - Cleaned")+
  guides(colour = guide_legend(override.aes = list(alpha = 1),reverse=TRUE))+
  scale_color_discrete(name="Daily Discharge Quantile", breaks = unique(SS_results$i),   labels=paste(Quantiles*100,"% - ",round(quantiles,0)," cfs"))+
  scale_x_datetime(date_labels = "%b\n%Y")+
  ylab("UV Discharge (cfs)")
```

#Specific Stage Result Plots

Now that outliers have been removed, we can print all our results. First in one plot for all quantiles. Next are individual plots for all target quantiles

## All Quantiles Plot
```{r, echo=FALSE,message=FALSE, fig.height=8, fig.width=12}
#Plot Specific Stage Results for All Quantiles
ggplot(data=subset(SS_results_clean), aes(x=dateTime,y=GH_Inst,color=as.factor(i)))+
  geom_point(alpha=0.3)+
  ggtitle(paste("USGS ", Site, sitedata$station_nm), subtitle = "Specific Stage for all Flow Quantiles")+
  theme_ss(14, "sans")+
  scale_color_discrete(name="Daily Discharge Quantile", breaks = unique(SS_results$i), labels=paste(Quantiles*100,"% - ",round(quantiles,0)," cfs"))+
  scale_x_datetime(date_labels = "%b\n%Y")+
  guides(colour = guide_legend(reverse=TRUE))+
  scale_x_datetime(date_labels = "%b\n%Y",date_breaks = "1 year")+
  ylab("Stage (ft)")+
  #geom_smooth(method="gam",alpha=0.5,se=FALSE)
  geom_smooth(alpha=0.05,se=FALSE)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2010-10-01")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2018-9-11")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2018-5-27")), linetype=2)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2016-7-30")), linetype=2)+
  theme(axis.title.x=element_blank())
```

## Target Quantile Plots
```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.height=3, fig.width=6}

for (v in seq(Quantiles)){
target_plot<-ggplot(data=subset(SS_results_clean,i==v), aes(x=dateTime,y=GH_Inst))+
  geom_smooth(alpha=0.3, color="grey", span=0.5, se=FALSE)+
  geom_point(alpha=0.3)+
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste("Specific Stage at", round(quantiles[[v]],2), " cfs, ", Quantiles[v]*100,"% of mean daily flow"))+
  theme_ss(14, "sans")+
  scale_color_discrete(name="Discharge\nQuantile", breaks = unique(SS_results$i), labels=paste(Quantiles*100,"% - ",round(quantiles,0)," cfs"))+
  guides(colour = guide_legend(reverse=TRUE))+
  scale_x_datetime(date_labels = "%b\n%Y",date_breaks = "1 year")+
  ylab(paste("Stage (ft)"))+
  #geom_smooth(method="gam",alpha=0.5,se=FALSE)
  geom_vline(xintercept=as.numeric(as.POSIXct("2010-10-01")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2018-9-11")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2018-5-27")), linetype=2)+
  geom_vline(xintercept=as.numeric(as.POSIXct("2016-7-30")), linetype=2)+
  theme(axis.title.x=element_blank())
print(target_plot)
}
```

# Let's explore site measurements (This is for all sites, not specific to the above analysis)

##Download site measurement data and derive secondary stats
```{r Download site measurement data and calculate derived stats}
Site <- Site
Sites <- c("01589000","01589025","01589035")

meas <- readNWISmeas(Sites, 
                     startDate = startDate,
                     endDate = "", 
                     tz = "America/New_York", 
                     expanded = TRUE, 
                     convertType = TRUE)

#meas
#area*velocity = discharge 
#area/width = height
#Fr = Velocity/sqrt(32.2*depth) <-- check that this is correct

meas <- meas %>%
  mutate(chan_depth = chan_area/chan_width) %>%
  mutate(Froude = chan_velocity/(sqrt(32.2*chan_depth)))
# Fr <1 - subcritical flow
# Fr >1 - supercritical flow
# Fr = critical flow
```

##Plot Estimated Froude number in time
```{r Plot Estimated Froude Number, echo=FALSE}
Q_window_low <- 500
Q_window_high <- 1000

meas %>%
  filter(discharge_va > 10 & discharge_va < 1200) %>%
ggplot(aes(x=measurement_dt, y=Froude, color=discharge_va)) +
  geom_point()+
  scale_color_viridis_c()  +
  ggtitle(paste("USGS ", "Patapsco River"),
          subtitle = paste0("Estimated Froude between ",Q_window_low,"-",Q_window_high," cfs"))+
  theme_ss(14, "sans")+
  geom_smooth(alpha=0.3, color="grey", span=0.2, se=FALSE)+
  xlim(as.Date("2009-10-01"),as.Date("2018-10-23"))+
  geom_vline(xintercept=as.numeric(as.Date("2010-10-01")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-9-11")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-5-27")), linetype=2)+
  geom_vline(xintercept=as.numeric(as.Date("2016-7-30")), linetype=2)+
  facet_grid(site_no~.)
```
##Plot estimated channel depth
```{r Measured Channel Depth, echo=FALSE}
Q_window_low <- 10
Q_window_high <- 100

meas %>%
  filter(discharge_va > Q_window_low & discharge_va < Q_window_high) %>%
ggplot(aes(x=measurement_dt, y=-chan_depth, color=discharge_va)) +
  geom_point()+
  scale_color_viridis_c()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste0("Channel Depth Measured between ",Q_window_low,"-",Q_window_high," cfs"))+
  theme_ss(14, "sans")+
  geom_smooth(alpha=0.3, color="grey", span=0.2, se=FALSE)+
    geom_vline(xintercept=as.numeric(as.Date("2010-10-01")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-9-11")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-5-27")), linetype=2)+
  geom_vline(xintercept=as.numeric(as.Date("2016-7-30")), linetype=2)+
  xlim(as.Date("2009-10-01"),as.Date("2018-10-23"))+
  facet_grid(site_no~.)
```
##Plot measured channel area
```{r Measured Channel Area, echo=FALSE}
Q_window_low <- 10
Q_window_high <- 1000
meas %>%
  filter(discharge_va > Q_window_low & discharge_va < Q_window_high) %>%
ggplot(aes(x=measurement_dt, y=chan_area, color=discharge_va)) +
  geom_point()+
  scale_color_viridis_c()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste0("Channel Area Measured between ",Q_window_low,"-",Q_window_high," cfs"))+
  theme_ss(14, "sans")+
  geom_smooth(alpha=0.3, color="grey", span=0.2, se=FALSE)+
    geom_vline(xintercept=as.numeric(as.Date("2010-10-01")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-9-11")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-5-27")), linetype=2)+
  geom_vline(xintercept=as.numeric(as.Date("2016-7-30")), linetype=2) +    
  xlim(as.Date("2009-10-01"),as.Date("2018-10-23"))+
  facet_grid(site_no~.)


```
##Plot measured channel velocity
```{r Measured Channel Velocity, echo=FALSE}
Q_window_low <- 50
Q_window_high <- 100
meas %>%
  filter(discharge_va > Q_window_low & discharge_va < Q_window_high) %>%
ggplot(aes(x=measurement_dt, y=chan_velocity, color=discharge_va)) +
  geom_smooth(span=0.4, se=FALSE, alpha=0.3, color="grey")+
  geom_point()+
  scale_color_viridis_c()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste0(" Channel Velocity Measured between ",Q_window_low,"-",Q_window_high," cfs"))+
  theme_ss(14, "sans")+
  geom_vline(xintercept=as.numeric(as.Date("2010-10-01")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-9-11")), linetype=4)+
  geom_vline(xintercept=as.numeric(as.Date("2018-5-27")), linetype=2)+
  geom_vline(xintercept=as.numeric(as.Date("2016-7-30")), linetype=2)+
  xlim(as.Date("2009-10-01"),as.Date("2018-10-23"))+
  facet_grid(site_no~.)
```
##Plot measured channel velocity~discharge
```{r Channel velocity ~ Discharge by Year, echo=FALSE}
meas %>%
  #filter(discharge_va > 1 & discharge_va < 2000) %>%
  filter(measurement_dt > "2009-10-01" & measurement_dt < "2018-10-23") %>%
  separate(measurement_dt, c("Year", "Month","Day")) %>%
ggplot(aes(x=log(discharge_va), y=chan_velocity, color=as.numeric(Year),group=as.numeric(Year))) +
  geom_point()+
  scale_color_viridis_c()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste("Channel velocity for discharge, by year"))+
  theme_ss(14, "sans")+
#coord_trans(x="log", y="log")+
  geom_smooth(se=FALSE, method = "lm") +
  facet_grid(site_no~.)
```
###Yearly Animation
```{r Channel velocity ~ Discharge by Year -- Animated, echo=FALSE}
p_load(gganimate)

meas %>%
  #filter(discharge_va > 1 & discharge_va < 2000) %>%
  filter(measurement_dt > "2009-10-01" & measurement_dt < "2018-10-23") %>%
  separate(measurement_dt, c("Year", "Month","Day")) %>%
ggplot(aes(x=log(discharge_va), y=chan_velocity, color=site_no,group=site_no)) +
  geom_point()+
  scale_color_viridis_d()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste("Channel Velocity for discharge, by year"))+
  theme_ss(14, "sans")+
  #coord_trans(x="log", y="log")+
  geom_smooth(se=FALSE, method = "lm")+
  labs(title = 'Year: {closest_state}', x = 'Discharge (cfs) in log units', y = 'Channel Velocity') +
  transition_states(Year, transition_length = 2, state_length = 3)+
  enter_fade() + 
  exit_fade() +
  ease_aes('linear')
```

##Plot froude~discharge+Year
```{r Froude for discharge, echo=FALSE}
plot <- meas %>%
  filter(measurement_dt > "2009-10-01" & measurement_dt < "2018-10-23") %>%
  filter(discharge_va > 1 & discharge_va < 2000) %>%
  separate(measurement_dt, c("Year", "Month","Day")) %>%
ggplot(aes(x=log(discharge_va), y=Froude, color=as.numeric(Year),group=as.numeric(Year))) +
  geom_point()+
  scale_color_viridis_c()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste("Froude numbers for discharge, by year"))+
  theme_ss(14, "sans")+
#coord_trans(x="log", y="log")+
geom_smooth(se=FALSE, method = "lm")+
    facet_grid(site_no~.)


```
###Yearly Animation
```{r Froude for discharge - Yearly animation, echo=FALSE}
#p_load(stringi, rlang, glue, scales)
#devtools::install_github("dgrtwo/gganimate", local=FALSE)
# p_load(rlang, httr)
# p_load(BiocManager)

# library(devtools)
#install_github("dgrtwo/gganimate", dependencies = TRUE)


data<-meas %>%
  filter(measurement_dt > "2009-10-01" & measurement_dt < "2018-10-23") %>%
  filter(discharge_va > 1 & discharge_va < 2000) %>%
  separate(measurement_dt, c("Year", "Month","Day"))

plot<-ggplot(data=data, aes(x=log(discharge_va), y=Froude, color=site_no,group=site_no, frame=Year)) +
  geom_point()+
  scale_color_viridis_d()  +
  ggtitle(paste("USGS ", Site, sitedata$station_nm),
          subtitle = paste("Froude numbers for discharge, by year"))+
  theme_ss(14, "sans")+
  #coord_trans(x="log", y="log")+
  geom_smooth(se=FALSE, method = "lm")

ggplotly(plot)

```
